---
title: "The Effects of Weather Events in the US on health"
author: "emilliman5"
date: "September 21, 2014"
output: html_document
---
#Synopsis


#Data Processing

Load library/packages.

```{r}

library(ggplot2)
library(reshape2)
library(R.utils)
library(plyr)

```

Download, decompress and read in data. Remove unneeded columns to free up memory.
```{r cache=TRUE}

download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile = "NOAA_storm.csv.bz2", method="curl")

bunzip2(filename = "NOAA_storm.csv.bz2", overwrite = TRUE)

NOAA<-read.csv("NOAA_storm.csv")

head(NOAA)
tail(NOAA)

StormNOAA<-NOAA[,c(2,7,8,23,24,25,26,27,28)]
rm(NOAA)
head(StormNOAA)
```

Create a smaller dataset for scripting to save on computational time
```{r}
StormNOAA<-StormNOAA[sample(nrow(StormNOAA), 100000),]
list(StormNOAA$EVTYPE)
StormNOAA$EVTYPE<-droplevels(StormNOAA$EVTYPE)
str(StormNOAA)
```

#Results

Question 1: Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

The effect of weather events on population health can be assessed by injuries and fatalites for each event. I first tabulated the mean and total number of health incidents (injuries and fatalities) per event type. I then plotted the top 10% (either mean or total) of events to assess the most severe weather event to human health.

```{r}

health<-cbind(aggregate(x = StormNOAA[,4]+StormNOAA[,5], by = list(StormNOAA$EVTYPE), FUN = mean, data = StormNOAA), aggregate(StormNOAA[,4]+StormNOAA[,5] ~ StormNOAA$EVTYPE, FUN = sum,data = StormNOAA, na.rm=TRUE))[,c(1,2,4)]

colnames(health)=c("EVTYPE","Average", "Total")

subset(health, Average>quantile(health$Average, c(.95)) | Total>quantile(health$Total, c(.95)))

topTenNOAA<-StormNOAA[StormNOAA$EVTYPE %in% health[health$Average>quantile(health$Average, c(.95))|health$Total>quantile(health$Total, c(.95)),1],]

topTenNOAA$EVTYPE<-droplevels(topTenNOAA$EVTYPE)

par(cex.lab=0.25)
boxplot(topTenNOAA$INJURIES+topTenNOAA$FATALITIES~topTenNOAA$EVTYPE)
```

Question 2: Across the United States, which types of events have the greatest economic consequences?

