---
title: "Motor Trends Report"
author: "emilliman4"
date: "01/24/2015"
output: pdf_document
---


```{r}
library(gplots)
data(mtcars)
summary(mtcars)
pairs(mtcars)
heatmap.2(cor(mtcars),)
```



```{r, echo=FALSE}
fit<-lm(mpg~., mtcars)
summary(fit)
influence.measures(fit)
```


```{r}
fit1<-lm(mpg~factor(am), data=mtcars)
summary(fit1)$coeff
fit2<-lm(mpg~factor(am)+wt, data=mtcars)
summary(fit2)$coeff
fit3<-lm(mpg~factor(am)+wt+factor(am)*wt, data=mtcars)
summary(fit3)$coeff
fit4<-lm(mpg~factor(am)+factor(vs)+wt+factor(am)*wt, mtcars)
summary(fit4)
fit5<-lm(mpg~factor(am)+factor(vs)+wt+drat+factor(am)*wt, mtcars)
summary(fit5)

anova(fit1,fit2,fit3,fit4,fit5)
```

Did the student interpret the coefficients correctly?

Did the student do some exploratory data analyses?

Did the student fit multiple models and detail their strategy for model selection?

Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?

Did the student do a residual plot and some diagnostics?

Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
